### 2018-02-27 proposal.tex
Proposal written and WIP references

\documentclass{scrartcl}

\usepackage[hidelinks]{hyperref}
\usepackage[none]{hyphenat}
\usepackage{setspace}
%\doublespace Plz no, I need to proofread this

\title{Which packet compression techniques for player synchronisation are ideal for improving speed and reducing bandwidth in large-scale multi-player games?}
\subtitle{COMP130 - Computing Architecture}
\date{\today}
\author{1707981}

\begin{document}
\maketitle
\pagenumbering{arabic}

\section{Proposal}
The exponential bandwidth consumption resulting by distributing player data in large-scale multi-player games is a significant challenge to game developers. Server maintenance costs are high as a consequence of high bandwidth use. Furthermore, a high percentage of UK homes run on a broadband speed at around 15Mbps, which roughly translates.

This study is interested in finding new ways, in many places drawing inspiration from the recent and highly successful temporal video compression techniques achieved as a reaction to the growth of streaming services such as YouTube.

To identify potential costs, this study will first aim to discover the presently known costs of game server maintenance. This will help us understand how much could be gained or lost from a reduction in network packet traffic, and thus how much time, money and effort a company may invest in combating the issue.

Following this, the study will attempt to uncover present weaknesses in common practices. This will share technical insight into the most significant bottlenecks common in today's online games. A particular genre of interest is the MMO genre, as well as the growing `battle royale', genre both of which can enjoy masses of players synchronised in a single world. These games are particularly vulnerable to exponential bandwidth congestion due to the player traffic sent from each player to each other player.

Finally, it will propose techniques to reduce packet consumption. This may include existing game networking techniques, image compression, and video compression or streaming strategies.

Video compression is a particular area of interest, owing to its temporal nature: frames are compressed not in absolute terms, but relative to the previous frame. This is not unlike the changes in player states as they progress through a game, although the specific types of data are quite different. Whether or not image data compression can be transferred to player data for any specific technique will be carefully considered throughout the essay.

The rapid advancement of the video compression domains could expose new opportunities in the game industry to efficiently compress and distribute update packets in previously unexplored ways. Should this happen, it may reveal the possibility for larger-scale persistent multi-player games to be created at a reasonable and sustainable cost for the interested game developer.

\bibliography{references} 
\bibliographystyle{ieeetr}

\end{document}

### 2018-02-27 proposal.tex
WIP new proposal files

@@
 
-\title{Which packet compression techniques for player synchronisation are ideal for improving speed and reducing bandwidth in large-scale multi-player games?}
+\title{Which object-oriented architectures could best support QOS-oriented map streaming in a multiplayer persistent C++ game?}
 \subtitle{COMP130 - Computing Architecture}
@@
 
+Games to lookup: Barotrauma, Doom 3?
+
 \section{Proposal}
-The exponential bandwidth consumption resulting by distributing player data in large-scale multi-player games is a significant challenge to game developers.
-Server maintenance costs are high as a consequence of high bandwidth use.
-Furthermore, a high percentage of UK homes run on a broadband speed at around 15Mbps, which roughly translates.
+Free-to-play multiplayer games enjoy a high popularity in poorer countries such as China, Brazil and Mexico, yet the lower broadband speeds in these areas can pose a particular challenge to games with persistent player-driven worlds.
+These types of worlds require the server to upload level data to each player periodically, such that they can see the current version of an area as they enter it.
 
-This study is interested in finding new ways, in many places drawing inspiration from the recent and highly successful temporal video compression techniques achieved as a reaction to the growth of streaming services such as YouTube.
+Unfortunately, broadband network speeds remain notoriously unreliable especially in developing countries.
+In order for games to be available to this wider market, they must utilise ways to reduce bandwidth while still providing enough data for the game to be playable.
 
-To identify potential costs, this study will first aim to discover the presently known costs of game server maintenance.
-This will help us understand how much could be gained or lost from a reduction in network packet traffic, and thus how much time, money and effort a company may invest in combating the issue.
+Quality-of-service encompasses the concept of restricting traffic in high bandwidth scenarios such that the data most useful to the player is prioritised.
+However, to utilise such a system, a game's internal architecture--including class structure, design patterns--must be prepared to function with incomplete data.
 
-Following this, the study will attempt to uncover present weaknesses in common practices.
-This will share technical insight into the most significant bottlenecks common in today's online games.
-A particular genre of interest is the MMO genre, as well as the growing `battle royale', genre both of which can enjoy masses of players synchronised in a single world.
-These games are particularly vulnerable to exponential bandwidth congestion due to the player traffic sent from each player to each other player.
-
-Finally, it will propose techniques to reduce packet consumption. This may include existing game networking techniques, image compression, and video compression or streaming strategies.
-
-Video compression is a particular area of interest, owing to its temporal nature: frames are compressed not in absolute terms, but relative to the previous frame.
-This is not unlike the changes in player states as they progress through a game, although the specific types of data are quite different.
-Whether or not image data compression can be transferred to player data for any specific technique will be carefully considered throughout the essay.
-
-The rapid advancement of the video compression domains could expose new opportunities in the game industry to efficiently compress and distribute update packets in previously unexplored ways.
-Should this happen, it may reveal the possibility for larger-scale persistent multi-player games to be created at a reasonable and sustainable cost for the interested game developer.
-
+This essay explores the application of design patterns to quality-of-service streaming in games.
+First, the significance and impact of low bandwidth on the player's end will be explored and challenged.
+Then, previous methods and architectures employed by companies in the past will be identified.
+This includes traditional single-player system architectures employed by successful games, followed by those employed by multiplayer games.
+A point of interest is whether there is an existing difference.
+Finally, this essay will compare the most popular architectures in general and attempt to theoretically deduce
+ which may be the most efficient
+\nocite{*}
 \bibliography{references} 

### 2018-03-02 proposal.tex
New proposal finished

@@
 
-\title{Which object-oriented architectures could best support QOS-oriented map streaming in a multiplayer persistent C++ game?}
+\title{Which C++ architecture(s) could best support QOS-oriented map streaming in a player-building-based multiplayer game?}
 \subtitle{COMP130 - Computing Architecture}
@@
 \section{Proposal}
-Free-to-play multiplayer games enjoy a high popularity in poorer countries such as China, Brazil and Mexico, yet the lower broadband speeds in these areas can pose a particular challenge to games with persistent player-driven worlds.
-These types of worlds require the server to upload level data to each player periodically, such that they can see the current version of an area as they enter it.
+Free-to-play multiplayer games enjoy a high popularity in poorer countries such as China, Brazil and Mexico, yet the lower broadband speeds in these areas can pose a particular challenge to games with persistent player-built worlds.
+This type of world requires the server to upload level data periodically to players in order for them to enjoy the latest version of an area as they enter it.
+The game may thus employ several different compression, prioritisation and/or data rate compensation techniques in order to achieve smooth gameplay, even on a reliable network.
 
-Unfortunately, broadband network speeds remain notoriously unreliable especially in developing countries.
-In order for games to be available to this wider market, they must utilise ways to reduce bandwidth while still providing enough data for the game to be playable.
+The asset download period is critical for players, as the data they receive determines which parts of the game they can play whilst the rest downloads in the background.
+Broadband speeds remain notoriously unreliable, especially in developing countries, so in order for games to be accessible to this wider market, they must reduce their bandwidth while still providing enough data to be playable during the download period.
+This includes reordering the data stream, with a priority on the player's best interests, such as their abilities first, in a concept more broadly known as quality-of-experience (QoE).
 
-Quality-of-service encompasses the concept of restricting traffic in high bandwidth scenarios such that the data most useful to the player is prioritised.
-However, to utilise such a system, a game's internal architecture--including class structure, design patterns--must be prepared to function with incomplete data.
+There are many known applications of QoE.
+However, this essay more specifically explores which applications of fundamental design patterns are most suitable in the context of an online game where players may build the environment, a la Minecraft.
+It covers three core areas:
 
-This essay explores the application of design patterns to quality-of-service streaming in games.
-First, the significance and impact of low bandwidth on the player's end will be explored and challenged.
-Then, previous methods and architectures employed by companies in the past will be identified.
-This includes traditional single-player system architectures employed by successful games, followed by those employed by multiplayer games.
-A point of interest is whether there is an existing difference.
-Finally, this essay will compare the most popular architectures in general and attempt to theoretically deduce
- which may be the most efficient
+\begin{itemize}
+\item To gain further insight into the weight of the issue, the significance and impact of low bandwidth on the player's end will be explored and challenged.
+Grey literature such as developer retrospectives may provide the most information on this topic.
+\item To discover potential architectural strengths and flaws, existing methods and architectures used by developers will be identified.
+As there may be some architectures with a wide potential of uses, traditional single-player system architectures employed by successful games may be explored in addition to those in multiplayer games.
+Any trend of difference between them is a point of interest.
+\item Finally, this essay will compare the most popular architectures in general, and attempt to deduce which may be the most efficient through an analysis of each area's noted flaws and strengths.
+While the broadness of potential solutions is implied, and perhaps a key part of the appeal of game programming as a whole, the essay will conclude with its own particular proposal for the hypothetical building game.
+\end{itemize}
 \nocite{*}

### 2018-03-12 draft.tex
WIP draft.pdf + Keeping up-to-date for PC work

\documentclass{scrartcl}

\usepackage[hidelinks]{hyperref}
\usepackage[none]{hyphenat}
\usepackage{setspace}
%\doublespace Plz no, I need to proofread this

\title{Which C++ architecture(s) could best support QoE-oriented map streaming in a player-building-based multiplayer game?}
\subtitle{COMP130 - Computing Architecture}
\date{\today}
\author{1707981}

\begin{document}
\maketitle
\pagenumbering{arabic}

\begin{abstract}
An exploration into QoE techniques and their compatibility with various softwrae architecture reveals something that isn't revealed until I finish this flipping essay. Also, QoS is linked to QoE is linked to QoP is linked to Soviet Russia.
\end{abstract}

\section{Proposal}

\section{Challenges of low bandwidth}

\section{Notes}
\subsection{China market findings}
China is in fact the most profitable games market as of 2017 \cite{https://technology.ihs.com/593405/china-largest-games-market-continues-its-growth-story}, yet the Internet speeds lag behind (figuratively speaking) at 7.6Mbps. Another country notorious for its online gaming community is Brazil, notorious for its unreliable Internet speeds, which averaged about the same--6.8Mbps--in 2017.\cite{https://www.akamai.com/us/en/multimedia/documents/state-of-the-internet/q1-2017-state-of-the-internet-connectivity-report.pdf}. While a marked improvement over time is evident (doubling since 2015 \cite{https://www.akamai.com/us/en/multimedia/documents/state-of-the-internet/akamai-state-of-the-internet-report-q1-2015.pdf}), to the extent that low-bandwidth games such as FPSes are comfortably playable, and the player reputation as in \cite{} is reduced.

This presents a heavy roadblock for MMO developers

Internet speeds in Brazil are notoriously low, which is remarkable for the fifth top source of DDoS attacks in Q1 201\cite{https://www.akamai.com/us/en/multimedia/documents/state-of-the-internet/q4-2017-state-of-the-internet-security-report.pdf}

\subsection{Link between computer specs and Internet speeds}

\nocite{*}
\bibliography{references} 
\bibliographystyle{ieeetr}

\end{document}

### 2018-03-13 presentation.tex
Presentation could be done, but it couldn't be fun

\documentclass{beamer}
\usetheme{Boadilla}
\usepackage{cite}

% Look at other themes?

\title{Architecting Effective Map Streaming in Low-Bandwidth Scenarios}
\subtitle{An inspiring and amazing research question that everyone wants to know, but was too afraid to write a 2000-word essay about.}
\author{Louis Foy}
\institute{Falmouth University}
\date{\today}

\begin{document}
\begin{frame}
	\titlepage
\end{frame}

\section{The Question}
\begin{frame}
	\frametitle{The Question}
	\begin{itemize}
		\item Which C++ architecture(s) could best support QoE-oriented map streaming in a player-building-based multiplayer game?
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{A scenario}
	\begin{itemize}
		\item My company wants to make a large-scale, multiplayer building game
		\item Players can build complex structures in real-time
		\item It\textquoteright s releasing across the world
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{A problem}
	\begin{itemize}
		\item Map data downloads take a lot of bandwidth
		\item Large markets, such as Egypt, Brazil, and China, suffer from slow broadband 
		\begin{itemize}
			\item Even though China is the world’s strongest gaming market! \cite{chinamarket}
		\end{itemize}
		\item This risks leaving players waiting for downloads
	\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Intro: QoE}
\begin{itemize}
	\item QoE means Quality of Experience
	\item Focuses on directing data flow to the user’s best interests first
	\begin{itemize}
		\item For example, downloading their character’s running animations, before downloading the distant city of Goradonia
	\end{itemize}
	\item Commonly used in video streaming to balance motion and detail
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{User experience != bitrate}
\begin{itemize}
\item User experience is more complex!
\item For example, evidence suggests users prefer a smooth low-quality video over higher-quality, inconsistent video \cite{qoesmallstudy} \cite{qoelargestudy}
\item The amount of action in a scene can affect their perception \cite{qoelargestudy}
\begin{itemize}
\item More action, less focus on detail. Less action, more focus. Could this apply to games as well?
\end{itemize}
\item In a nutshell, intelligent prioritising of streams is key
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Prerequisites: Map streaming}
\begin{itemize}
	\item To identify ideal architectures for implementing QoE map streaming, let’s explore the necessary ingredients
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Prerequisites - Baseline}
\begin{itemize}
\item Map data uploader
\begin{itemize}
	\item Sends map data to clients
\end{itemize}
\item Map data downloader
\begin{itemize}
	\item Takes map data from server
	\item Must tolerate incomplete data
\end{itemize}
\item Map constructor
\begin{itemize}
	\item Builds the map in real-time
	\item May include collision generation, textures, and LoD
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Prerequisites - With QoE}
\begin{itemize}
	\item Map data prioritiser
	\begin{itemize}
		\item Decides which map data each player needs soonest
		\item Must work with data uploader
	\end{itemize}
	\item Map data compressors
	\begin{itemize}
		\item Reduces bandwidth by brute force
		\item Adds processing time on downloads: may introduce FPS lag
	\end{itemize}
	\item `Network LOD' generators
	\begin{itemize}
		\item Reduces details on distant player-made areas
		\item Different to regular LOD--regular LOD is predefined on your local PC
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Preliminary architectural theories}
\begin{itemize}
\item Multi-threading
\begin{itemize}
	\item Could reduce distressing lag during downloads
\end{itemize}
\item Adapters?
\begin{itemize}
	\item Could wrap compressed map data into usable geometry
\end{itemize}
\item Observers?
\begin{itemize}
	\item Could live in the map class, awaiting incoming map data
\end{itemize}
\item Or mediators?
\begin{itemize}
	\item Alternative to data transferral between networked map data and geometry
	\item Which one is better?
\end{itemize}
\item OOP?
\begin{itemize}
	\item OOP principles support all the above design patterns quite well
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{More to come}
\begin{itemize}
	\item More specific design patterns will be isolated pending further research into existing game titles
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Any questions?}
\begin{itemize}
	\item References:
\end{itemize}
\bibliographystyle{ieeetr}
\bibliography{references_presentation}
\end{frame}

\end{document}

### 2018-03-20 draft.tex
This'll have to do for now

@@
 
-\title{Which C++ architecture(s) could best support QoE-oriented map streaming in a player-building-based multiplayer game?}
+\usepackage{graphicx}
+\usepackage{float}
+\graphicspath{{images/}}
+
+\title{What C++ architecture could best support QoE-oriented map streaming in a player-building-based multiplayer game?}
 \subtitle{COMP130 - Computing Architecture}
@@
 \begin{abstract}
-An exploration into QoE techniques and their compatibility with various softwrae architecture reveals something that isn't revealed until I finish this flipping essay.
-Also, QoS is linked to QoE is linked to QoP is linked to Soviet Russia.
+Low broadband speeds in large gaming markets such as China and Brazil could hinder the advancement of multiplayer gaming, particularly for traffic-hungry projects such as MMOs or building games.
+From a creative standpoint, player experience is decidedly the most important factor in a game's design.
+Low broadband bandwidth increases download speeds and strips players of valuable gaming time, causing frustration.
+To aid this, Quality-of-Experience (QoE) techniques are proposed to mitigate the negative effects of limited bandwidth on the player experience, in a way that could be adapted to existing games with minimal difficulty.
 \end{abstract}
 
+\section{Introduction}
+\begin{figure}
+\includegraphics[width=0.75\linewidth]{Newzoo_2017_Global_Games_Market.png}
+\caption{The games market in 2017, according to NewZoo}
+\label{fig:market}
+\end{figure}
+The games market is continuing to grow at an immense rate, especially in developing countries. In East Asia, a growth of nearly 15\% over the last year has been observed.\cite{fig:market}
+
+This presents exciting creative opportunities for game developers to expand into new territories.
+However, a lag in technological advancement, particularly broadband, raises reasonable concerns of whether they can fully exploit these opportunities.
+The potential of games where players can build the map, such as the extremely popular (as suggested by Google Trends\cite{minecraftnite}) Minecraft and Fortnite, may be hindered by a lack of bandwidth necessary to download player-built resources at an acceptable speed.
+This may frustrate players by rendering them unable to access areas of the game in a timely manner.
+
+This study aims to discover the extent of this problem, followed by proposals of architectural methods to implement systems that could mitigate the negative effects on the player experience in an unreliable networking environment: This largely comprises of QoE (quality-of-experience) improvements.
+The programming language in question is C++, while the hypothetical game assumed a 3D environment wherein the player can manipulate the terrain and build meshes.
+The goal is to propose a reasonable set of design patterns to be used for QoE-oriented map streaming in a large-scale game, and how they could be applied to an existing game engine.
+
+This analysis will take place over the following steps:
+\begin{itemize}
+\item Exploration into the games market, and its broadband speeds
+\item A brief introduction of player experience as a separate concept to content quality
+\item Existing strategies used by games for a) map streaming and b) online synchronisation, and
+\item A proposed technical outline of a system that could facilitate appropriate level streaming considering the aforementioned variables.
+\end{itemize}
+
+\section{The wider market}
+China is in fact the most profitable games market as of 2017 \cite{chinamarket}.
+Furthermore, in March 2018, 8 of the 10 most popular Steam games in China were partly or fully multiplayer \cite{steamchina}.
+This suggests China is a valuable target for online games, yet its Internet speeds lag behind (figuratively speaking) the world, at an average of 7.6Mbps \cite{webspeeds}.
+
+Another country notorious for its online gaming community is Brazil, whose Internet speeds averaged about the same--6.8Mbps--in 2017.\cite{webspeeds}.
+While a marked improvement is evident (doubling since 2015 \cite{webspeeds2015}), to the extent that low-bandwidth games such as FPSes are comfortably playable [cite?], this low speed closes doors to potential large-scale player-built online games.
+
+(Internet speeds in Brazil are notoriously low--an unfortunate feat for the fifth top source of DDoS attacks in early 2017\cite{websecurity}--)
+
+\section{Player experience}
+Video streaming is a heavily researched area; perhaps more than games, owing to greater dedicated R\&D toward streaming services such as the Amazon FireTV[cite?].
+Considering the aesthetic link between video quality and quality of gameplay--including pixel resolution, lag spike rate, and frame rate--we can draw some preliminary conclusions about a player's interests from studies.
+
+It has been documented \cite{qoelargestudy} that user enjoyment is greater in a scenario of low-detail, constant frame streaming, than one of higher-detail, inconsistent frame streaming.
+In other words, the user is not interested in the rate of high-detailed data--they are interested in enjoying a smoother experience.
+
+A smooth experience is necessary to retain game players.
+Studies have noted latency \cite{qossensitivity} \cite{lagragequits}, jitter and packet loss \cite{lagragequits} to all have negative impacts, sometimes leading to quitting altogether.
+While this is not widely studied in building games--but FPSes and MMOs--games such as the hugely popular (cite?) Fortnite, which features building elements with shooting gameplay, suggest that there may be a demand for player building in those contexts regardless.
+The potential loss of players therefore justifies the development of strategies to improve the player experience on an unreliable network, especially considering the high traffic potential of level streaming.
+
+\section{Existing strategies}
+Granted, it is fair to say player experience-driven streaming already exists, though this assumes a higher data rate, and takes the form of open-world streaming from disc.
+Engines such as the Unreal engine provide World Composition, allowing the world to be partly loaded in small sections as the player approaches them \cite{unrealcomposition}.
+This is characterised by the loading of reduced LOD (Level-of-Detail) versions of areas distant to the camera, and the high LOD versions of close areas.
+This hides detail in the distance at the benefit of reducing the memory footprint and rendering times: both a potential improvement to the asset download time and frame rate, and consequently an improvement to the player's perception of the game.
+
+This streaming method provides data to the player in a spatially-aware manner such that the player has fastest access to the area immediately around them.
+However, the key difference in the online equivalent system is the data rate.
+In the PS4 and Xbox One's case, static map streaming from the disc this defaults at 27MB/s (Cite this)--a significantly higher rate than China's average Internet speed of 7.6Mbps, or 0.95 MB/s.
+At only 4\% of the typical disc data rate, it is reasonable to question whether game level streaming is prepared for such a bottleneck.
+
 \section{Proposal}
+Due to an extreme lack of time and an extremely tired caffeine-free (and vitamin-deficient) author, a tacked-together solution is hastily proposed by this essay draft.
 
-\section{Challenges of low bandwidth}
+\begin{figure}
+	\includegraphics[width=0.7\linewidth]{Basic_Map_Streamer}
+	\caption{A class diagram demonstrating a map streaming component connected to a basic Map class.}
+	\label{fig:simplesystem}
+\end{figure}
 
-\section{Notes}
-\subsection{China market findings}
-China is in fact the most profitable games market as of 2017 \cite{https://technology.ihs.com/593405/china-largest-games-market-continues-its-growth-story}, yet the Internet speeds lag behind (figuratively speaking) at 7.6Mbps.
-Another country notorious for its online gaming community is Brazil, notorious for its unreliable Internet speeds, which averaged about the same--6.8Mbps--in 2017.\cite{https://www.akamai.com/us/en/multimedia/documents/state-of-the-internet/q1-2017-state-of-the-internet-connectivity-report.pdf}.
-While a marked improvement over time is evident (doubling since 2015 \cite{https://www.akamai.com/us/en/multimedia/documents/state-of-the-internet/akamai-state-of-the-internet-report-q1-2015.pdf}), to the extent that low-bandwidth games such as FPSes are comfortably playable, and the player reputation as in \cite{} is reduced.
+The class diagram above depicts the client end of the map streaming game component.
+As some partly-downloaded data is not needed by the game component until it's complete, the process is divided into two concurrent, non-conflicting threads.
+Flow takes place in the following fashion:
 
-This presents a heavy roadblock for MMO developers
+\begin{figure}[H]
+	\includegraphics[width=0.7\linewidth]{Basic_Flowchart.png}
+	\caption{Process flow of the two threads in parallel}
+	\label{fig:simpleflow}
+\end{figure}
 
-Internet speeds in Brazil are notoriously low, which is remarkable for the fifth top source of DDoS attacks in Q1 201\cite{https://www.akamai.com/us/en/multimedia/documents/state-of-the-internet/q4-2017-state-of-the-internet-security-report.pdf}
+In this example, the client begins by sending the player and camera location as soon as possible.
+Sometimes these two variables may be different: it is important that both areas are prepared for the player's viewing.
+The server is assumed to prioritise the level chunk geometry from closest to furthest.
 
-\subsection{Link between computer specs and Internet speeds}
+A \textbf{level chunk} is defined as a region of space where static level geometry resides.
+The shape of a level chunk varies across games, though in a building game, a uniform shape could be chosen such that there are no notable inefficiencies to building in one place over another.
+Existing games use various strategies.
+For example, Sunset Overdrive (Insomniac Games, 2014) used hexagonal chunks by design and radial chunks in practice.
+This was optimised for the player's freedom of movement.
+However, an earlier game by the same company, Fuse (Insomniac Games, 2013), used rectangular chunks.
+This was optimised for linear hall-based gameplay.
+(CITE ALL THIS).
 
-\nocite{*}
+The \textbf{level streamer} is responsible for collecting \textbf{Chunk Packs} (partial, LOD-specific level chunk data downloaded from the server) and transferring them to the physical \textbf{Map} when it deems appropriate.
+In the above example, it does so as soon as it is immediately downloaded.
+However, it may benefit the player experience to record the amount of processing time spent during recent frames, and responding by slowing the conversion process or postponing it.
+This could help preserve a high frame rate and provide a smoother experience.
+
+\section{Total Conclusion}
+There are further systems to be explored, citations to be added and slightly passive-aggressive expressions of frustration to be passive-aggressively expressed at the recent surge in workload that are beyond the scope of this essay draft, and the author's physical capabilities at 3am, let alone intellectual ones.
+The author as such has little idea what they are doing and submits this draft to the peer review with great hesitation and a tear in his eye.
+
 \bibliography{references} 

### 2018-03-20 draft.tex
Updated and working through self-review. Let's recover some of those lost marks!!

@@
 \graphicspath{{images/}}
+
+\newcommand{\source}[1]{\caption*{Source: {#1}} }
+% Above code sourced from Xavi, Stack Overflow, https://tex.stackexchange.com/questions/95029/add-source-to-figure-caption @ 20/03/2018
 
@@
 \section{Introduction}
-\begin{figure}
+The games market is continuing to grow at an immense rate, especially in developing countries.
+In East Asia, a growth of nearly 15\% over the last year has been observed, as shown in Fig.
+\ref{fig:market}.
+
+\begin{figure}[H]
 	\includegraphics[width=0.75\linewidth]{Newzoo_2017_Global_Games_Market.png}
-\caption{The games market in 2017, according to NewZoo}
+	\caption{The games market in 2017, according to NewZoo (Source: \cite{globalmarketpic})}
 	\label{fig:market}
 \end{figure}
-The games market is continuing to grow at an immense rate, especially in developing countries. In East Asia, a growth of nearly 15\% over the last year has been observed.\cite{fig:market}
 
-This presents exciting creative opportunities for game developers to expand into new territories.
+This presents profitable opportunities for game developers to expand into new territories.
 However, a lag in technological advancement, particularly broadband, raises reasonable concerns of whether they can fully exploit these opportunities.
-The potential of games where players can build the map, such as the extremely popular (as suggested by Google Trends\cite{minecraftnite}) Minecraft and Fortnite, may be hindered by a lack of bandwidth necessary to download player-built resources at an acceptable speed.
-This may frustrate players by rendering them unable to access areas of the game in a timely manner.
+The potential of games where players can build the map, such as the extremely popular (as suggested by Google Trends \cite{minecraftnite}) Minecraft and Fortnite, may be hindered by a lack of bandwidth necessary to download player-built resources at an acceptable speed.
+A slow download of map resources may frustrate players by rendering them unable to access areas of the game in a timely manner.
 
-This study aims to discover the extent of this problem, followed by proposals of architectural methods to implement systems that could mitigate the negative effects on the player experience in an unreliable networking environment: This largely comprises of QoE (quality-of-experience) improvements.
+This study aims to discover the extent of this problem, followed by proposals of architectural methods to implement systems that could mitigate the negative effects of an unrealiable networking environment on the player experience.
+These can be described as QoE (quality-of-experience) improvements where the bandwidth is not improved, but the specific targeting of resources is.
 The programming language in question is C++, while the hypothetical game assumed a 3D environment wherein the player can manipulate the terrain and build meshes.
@@
 
-\section{The wider market}
-China is in fact the most profitable games market as of 2017 \cite{chinamarket}.
-Furthermore, in March 2018, 8 of the 10 most popular Steam games in China were partly or fully multiplayer \cite{steamchina}.
-This suggests China is a valuable target for online games, yet its Internet speeds lag behind (figuratively speaking) the world, at an average of 7.6Mbps \cite{webspeeds}.
+\section{The Wider, Slower Market}
+China is the most profitable games market as of 2017 \cite{chinamarket}.
+Furthermore, according to SteamSpy in March 2018, 8 of the 10 most popular Steam games in China were partly or fully multiplayer \cite{steamchina}.
+This suggests China is a valuable target for online games.
+However, compared to most developed countries, China's Internet speeds lag behind (figuratively speaking), at an average of 7.6Mbps \cite{webspeeds}.
+This inspires concerns that the potential of the multi-player gaming industry, particularly in asset-heavy games like building games, could be hindered in this potentially profitable region.
 
-Another country notorious for its online gaming community is Brazil, whose Internet speeds averaged about the same--6.8Mbps--in 2017.\cite{webspeeds}.
-While a marked improvement is evident (doubling since 2015 \cite{webspeeds2015}), to the extent that low-bandwidth games such as FPSes are comfortably playable [cite?], this low speed closes doors to potential large-scale player-built online games.
+Similarly, another country notorious for its online gaming community is Brazil.
+According to SteamSpy, 9 of the top 10 games had a multiplayer component, with 8 boasting primarily multiplayer gameplay \cite{steambrazil}.
+Brazil's Internet speed average about the same--6.8Mbps--in 2017 \cite{webspeeds}.
+While a marked improvement over time is evident (doubling since 2015 \cite{webspeeds2015}), to the extent that low-bandwidth games such as FPSes are comfortably playable [cite?], it puts to question whether the development of large-scale player-built online games is halted by this lack of gross speed (an ironic failing for the world's fifth top source of DDoS attacks in early 2017 \cite{websecurity}).
 
-(Internet speeds in Brazil are notoriously low--an unfortunate feat for the fifth top source of DDoS attacks in early 2017\cite{websecurity}--)
+There is a lack of definitive research of the impact of low broadband speed in games.
+Instead, many studies such as [CITE] are focused on reducing and optimising traffic.
+This is helpful to developers, but lacks a point of reference or scale, making it plausible that a developer could fear the unknown bottlenecks of slow broadband.
+Until this is further understood, it could be thus beneficial to design a system around scalable broadband speed, such that it can guarantee an acceptably 'fun' quality even in extremely slow network conditions.
+However, the definition of 'fun' may be at odds with the industry's tendency to create games that are bigger, more detailed, and overall more data-heavy as they evolve.
+(CITE) This prompts an exploration into what is 'fun', in a term more aptly described as quality of experience.
 
-\section{Player experience}
+\section{Quality of player experience}
 Video streaming is a heavily researched area; perhaps more than games, owing to greater dedicated R\&D toward streaming services such as the Amazon FireTV[cite?].
-Considering the aesthetic link between video quality and quality of gameplay--including pixel resolution, lag spike rate, and frame rate--we can draw some preliminary conclusions about a player's interests from studies.
+Considering the link between video quality and quality of gameplay aesthetic--including pixel resolution, lag spike rate, and frame rate--some preliminary conclusions about a player's interests can be drawn from this research.
 
@@
 
-A smooth experience is necessary to retain game players.
+This smooth experience is also necessary to retain game players.
 Studies have noted latency \cite{qossensitivity} \cite{lagragequits}, jitter and packet loss \cite{lagragequits} to all have negative impacts, sometimes leading to quitting altogether.
-While this is not widely studied in building games--but FPSes and MMOs--games such as the hugely popular (cite?) Fortnite, which features building elements with shooting gameplay, suggest that there may be a demand for player building in those contexts regardless.
-The potential loss of players therefore justifies the development of strategies to improve the player experience on an unreliable network, especially considering the high traffic potential of level streaming.
+While this is not widely studied in building games--but FPSes and MMOs--games such as the recent and hugely popular (cite?) Fortnite, which feature both building elements and shooting gameplay, suggest that there may be new demand for player building in those contexts.
+The potential loss of players therefore justifies the development of strategies to improve the player experience on an unreliable network, especially considering the high traffic potential of level streaming (CITE THIS!).
+
+\section{Streaming costs}
+A hypothetical budget map of descending streaming costs follows:
+- Textures
+- Leel geometry
+- Player animations
+- LOOK AT A GAME'S MEMORY BUDGET
 

### 2018-03-23 draft.tex
Updated and committing while the Existing Strategies section is still Existing.

@@
 
-\title{What C++ architecture could best support QoE-oriented map streaming in a player-building-based multiplayer game?}
+\title{What C++ architecture could best support Quality-of-Experience-oriented map streaming in a multiplayer building game?}
 \subtitle{COMP130 - Computing Architecture}
@@
 \begin{abstract}
-Low broadband speeds in large gaming markets such as China and Brazil could hinder the advancement of multiplayer gaming, particularly for traffic-hungry projects such as MMOs or building games.
+Low broadband speeds across the world could hinder the advancement of multiplayer gaming, particularly for traffic-hungry projects such as MMOs or building games.
 From a creative standpoint, player experience is decidedly the most important factor in a game's design.
-Low broadband bandwidth increases download speeds and strips players of valuable gaming time, causing frustration.
+When downloading the game environment, low bandwidth increases download speeds and temporarily strips players of their autonomy, causing frustration.
 To aid this, Quality-of-Experience (QoE) techniques are proposed to mitigate the negative effects of limited bandwidth on the player experience, in a way that could be adapted to existing games with minimal difficulty.
@@
 \begin{figure}[H]
+	\centering
 	\includegraphics[width=0.75\linewidth]{Newzoo_2017_Global_Games_Market.png}
@@
 
-This study aims to discover the extent of this problem, followed by proposals of architectural methods to implement systems that could mitigate the negative effects of an unrealiable networking environment on the player experience.
-These can be described as QoE (quality-of-experience) improvements where the bandwidth is not improved, but the specific targeting of resources is.
+This study aims to discover the extent of this problem, followed by proposals of architectural methods to implement systems that could mitigate the negative effects of an unreliable networking environment on the player experience.
+These can be described as QoE (Quality-of-Experience) improvements where the bandwidth is not improved, but instead coordinated where the player needs it most.
 The programming language in question is C++, while the hypothetical game assumed a 3D environment wherein the player can manipulate the terrain and build meshes.
@@
 This analysis will take place over the following steps:
+
 \begin{itemize}
-\item Exploration into the games market, and its broadband speeds
-\item A brief introduction of player experience as a separate concept to content quality
-\item Existing strategies used by games for a) map streaming and b) online synchronisation, and
-\item A proposed technical outline of a system that could facilitate appropriate level streaming considering the aforementioned variables.
+\item Exploration into the games market and its broadband speeds
+\item A brief introduction to player experience as a separate concept to content quality
+\item Existing strategies used by games for map streaming
+\item A proposed technical outline of a system that could facilitate level streaming with priority on player experience
 \end{itemize}
 
-\section{The Wider, Slower Market}
+\section{The Big, Slow Market}
 China is the most profitable games market as of 2017 \cite{chinamarket}.
@@
 However, compared to most developed countries, China's Internet speeds lag behind (figuratively speaking), at an average of 7.6Mbps \cite{webspeeds}.
-This inspires concerns that the potential of the multi-player gaming industry, particularly in asset-heavy games like building games, could be hindered in this potentially profitable region.
 
@@
 According to SteamSpy, 9 of the top 10 games had a multiplayer component, with 8 boasting primarily multiplayer gameplay \cite{steambrazil}.
-Brazil's Internet speed average about the same--6.8Mbps--in 2017 \cite{webspeeds}.
-While a marked improvement over time is evident (doubling since 2015 \cite{webspeeds2015}), to the extent that low-bandwidth games such as FPSes are comfortably playable [cite?], it puts to question whether the development of large-scale player-built online games is halted by this lack of gross speed (an ironic failing for the world's fifth top source of DDoS attacks in early 2017 \cite{websecurity}).
+Brazil's Internet speed averaged about the same--6.8Mbps--in 2017 \cite{webspeeds} (an ironic failing for the world's fifth top source of DDoS attacks in early 2017 \cite{websecurity}).
 
-There is a lack of definitive research of the impact of low broadband speed in games.
+Speeds in the UK and US meanwhile average at over twice as much, 16.9Mbps and 18.7Mbps respectively.
+While this study was originally expecting a bigger discrepancy, it could surmise that the map streaming engines of large-scale player-built online games will require optimisation to compensate the lack of data bandwidth compared to, for example, disk speeds.
+In the Sony PS4's case, streaming speed from the disc maxes out at 27MB/s (at a 1x speed of 36Mbps \cite{bluray} multiplied by 6 \cite{ps4specs})--a significantly higher rate than Brazil's average Internet speed of 6.8Mbps, or 0.85 MB/s.
+This is only 3\% of the typical disc data rate.
+
+There is a lack of definitive research on the impact of low broadband speed in map streaming.
 Instead, many studies such as [CITE] are focused on reducing and optimising traffic.
-This is helpful to developers, but lacks a point of reference or scale, making it plausible that a developer could fear the unknown bottlenecks of slow broadband.
-Until this is further understood, it could be thus beneficial to design a system around scalable broadband speed, such that it can guarantee an acceptably 'fun' quality even in extremely slow network conditions.
-However, the definition of 'fun' may be at odds with the industry's tendency to create games that are bigger, more detailed, and overall more data-heavy as they evolve.
-(CITE) This prompts an exploration into what is 'fun', in a term more aptly described as quality of experience.
+This is helpful to developers, but lacks a point of reference or scale in terms of user experience, making the actual negative impacts of low bandwidth relatively uncertain.
+Until this is further understood, it could be beneficial to design a system around scalable broadband speed, such that it can guarantee an acceptably 'fun' quality even in extremely slow network conditions.
+However, the definition of 'fun' may be at odds with the industry's tendency to create games that are bigger, more detailed, and overall more data-heavy as they evolve \cite{graphicsvsexperience}.
+This prompts an exploration into what is 'fun', in a term more aptly described as quality of experience.
 
-\section{Quality of player experience}
-Video streaming is a heavily researched area; perhaps more than games, owing to greater dedicated R\&D toward streaming services such as the Amazon FireTV[cite?].
-Considering the link between video quality and quality of gameplay aesthetic--including pixel resolution, lag spike rate, and frame rate--some preliminary conclusions about a player's interests can be drawn from this research.
+\section{Quality of player experience} \label{player}
+Video streaming is a heavily researched area; perhaps more than games, owing to significant dedicated R\&D toward streaming services such as the Amazon FireTV \cite{amazonresearch}.
+Yet considering the link between video quality and quality of gameplay aesthetic--including pixel resolution, lag spike rate, and frame rate--the basic principles and goals could perhaps apply to games as well.
 
-It has been documented \cite{qoelargestudy} that user enjoyment is greater in a scenario of low-detail, constant frame streaming, than one of higher-detail, inconsistent frame streaming.
-In other words, the user is not interested in the rate of high-detailed data--they are interested in enjoying a smoother experience.
+In the context of video, it has been documented \cite{qoelargestudy} that user enjoyment is greater in a scenario of low-detail, constant frame streaming, than one of higher-detail, inconsistent frame streaming.
+In other words, the user is not interested in the rate of high-detailed data, but in enjoying a smoother experience.
 
-This smooth experience is also necessary to retain game players.
+In the context of games, this smooth experience is also shown to be necessary to retain game players.
 Studies have noted latency \cite{qossensitivity} \cite{lagragequits}, jitter and packet loss \cite{lagragequits} to all have negative impacts, sometimes leading to quitting altogether.
-While this is not widely studied in building games--but FPSes and MMOs--games such as the recent and hugely popular (cite?) Fortnite, which feature both building elements and shooting gameplay, suggest that there may be new demand for player building in those contexts.
-The potential loss of players therefore justifies the development of strategies to improve the player experience on an unreliable network, especially considering the high traffic potential of level streaming (CITE THIS!).
+While this is not widely studied in building games--but FPSes and MMOs--games such as the recent and highly popular \cite{topgames} Fortnite, which features both building elements and shooting gameplay, demonstrates that the two are not necessarily mutually exclusive.
 
-\section{Streaming costs}
-A hypothetical budget map of descending streaming costs follows:
-- Textures
-- Leel geometry
-- Player animations
-- LOOK AT A GAME'S MEMORY BUDGET
+Another study \cite{graphicsvsexperience} noted the importance of animations in a game.
+Similarly, this centres around the concept of motion, and as above, suggests that graphical detail is of smaller importance than the player's ability to gain immediate feedback.
+This is supported at a base psychological level by R.
+Ryan and E.
+Deci's reputable research on human motivation \cite{motivation} (especially reputable in that they have disproportionately few people to cite, other than themselves, throughout their entire paper), which notes a human desire to explore, apply their skills, and learn from the feedback.
+
+These findings place high importance on the motion, control, and abilities of the player rather than visual fidelity of the background.
+In map streaming, this translates to the importance of giving the player base landscapesto move within and toward.
+This places the level mesh, and perhaps interactive objects (noting the need for motion), above textures and sounds in the priority list.
+
+\section{Quality of experience -- Priorities}
+To maximise player experience, a preliminary priority list is thus proposed:
+
+\begin{itemize}
+	\item Player movement
+	\item Player avatar
+	\item Textureless nearby level geometry
+	\item Interactive objects
+	\item Player animations
+	\item Object animations
+	\item Textures
+	\item Miscellaneous
+\end{itemize}
+
+However, this is not an exact science.
+For example, level textures may come in a low-quality form before their final form.
+This would show evidence of loading progress to the player.
+While it could be argued as invasive, a player would likely deem this more favourable than--for example--a loading screen, as observed in software studies \cite{loadingscreens}.
+Furthermore, level textures might not necessarily be streamed, unless the players should be able to draw their own.
+
+While they are described above, the specifics of animation, player model, object and texture streaming are beyond the scope of this essay.
+They could, however, very likely be adapted in a similar fashion.
 
 \section{Existing strategies}
-Granted, it is fair to say player experience-driven streaming already exists, though this assumes a higher data rate, and takes the form of open-world streaming from disc.
-Engines such as the Unreal engine provide World Composition, allowing the world to be partly loaded in small sections as the player approaches them \cite{unrealcomposition}.
-This is characterised by the loading of reduced LOD (Level-of-Detail) versions of areas distant to the camera, and the high LOD versions of close areas.
-This hides detail in the distance at the benefit of reducing the memory footprint and rendering times: both a potential improvement to the asset download time and frame rate, and consequently an improvement to the player's perception of the game.
+Locality-based streaming prioritisation already exists, and is used in open-world games to save memory and loading times.
+Engines such as the Unreal Engine provide World Composition, allowing the world to be partly loaded in small sections as the player approaches them \cite{unrealcomposition}.
+This is characterised by the loading of low-LOD (Level-of-Detail) versions of level components distant to the camera, and the high-LOD versions of close ones, all in a background process.
+This hides level details in the distance at the benefit of reducing the memory footprint, hiding loading times, and increasing the frame rate.
+The latter two in particular improves the player's perception of the game motion, and provides data to the player in a spatially-aware manner such that the player has immediate access to the areas closest to them.
 
-This streaming method provides data to the player in a spatially-aware manner such that the player has fastest access to the area immediately around them.
-However, the key difference in the online equivalent system is the data rate.
-In the PS4 and Xbox One's case, static map streaming from the disc this defaults at 27MB/s (Cite this)--a significantly higher rate than China's average Internet speed of 7.6Mbps, or 0.95 MB/s.
-At only 4\% of the typical disc data rate, it is reasonable to question whether game level streaming is prepared for such a bottleneck.
+However, systems such as the above assume a higher data rate, and is typically sourced from a disk or hard disk.
+
+\section{Proposal}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.7\linewidth]{Server_Side_Streamer.png}
+	\caption{Partial class diagram illustrating the potential map streaming server},
+	\label{fig:serversystem}
+\end{figure}
+
+The proposed system would comprise the following key classes:
+
+\textbf{Map}: Base map class.
+This is shared between the client and server.
+By partitioning the map into several 'chunks', separate pieces of it can be downloaded at a time, allowing the player to roam incomplete areas in short notice.
+\textbf{ChunkDestination}: A collection of \textbf{ChunkPacks} that may be implemented in multiple ways.
+In this case, it is connected to \textbf{MessageSender} for immediate transmission.
+In more advanced cases, it could also, at a later stage, be implemented on the client machine as a map storage manager, which could cache map data on a local machine for later use.
+\textbf{MapUpstreamer}: The MapUpstreamer is effectively a \textit{Mediator} (described as a class that manages the collaboration between objects \cite{designpatterns}) technically owned by neither Map nor ChunkDestination.
+This method was chosen so that MapUpstreamer could, if necessary, operate independently in parallel to the other tasks.
+An observer pattern could reasonably be used, but this typically means synchronising the data on an immediate basis \cite{designpatterns}, something which is not necessary for the time-sensitive task of prioritising the data flow.
+\textbf{BandwidthMonitor}: To achieve data rate consistency as likely desired by players (see Section \ref{player}), this class monitors bandwidth over time and adjusts the data rate to the most consistent possible.
+The specifics of this system, better described as bandwidth-smoothing, are beyond the scope of this essay.
+A broad comparison of bandwidth-smoothing techniques as illustrated by W.
+Feng and J.
+Rexford is recommended for further reading \cite{bandwidthsmoothing}.
+\textbf{ChunkPack}: This interface stores components of map data in various possible class-specific ways, and provides the function \textit{generateChunk} such that it can inject itself into the Map.
+An example of a subclass--\textbf{ChunkPack_Delta}--is demonstrated in Fig.
+\ref{fig:clientsystem}.
+
+The client side of the system we propose adopts the following hierarchy.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.7\linewidth]{Client_Side_Streamer.png}
+	\caption{Partial class diagram illustrating the potential map streaming client},
+	\label{fig:clientsystem}
+\end{figure}
+
+- The Map Streamer is a Listener class.
+This is to enable it to receive data from any input.
+A particular benefit of this approach is the ability to add different sources of data; for example, a local cache so that a player can return to a region without redownloading it.
+- The \textbf{ChunkPack} is an interface that can manipulate map chunks by inserting or removing data of its choice.
+By leaving this responsibility to the ChunkPack class, subclasses such as \textbf{ChunkPackDelta} can be created to manipulate the chunk data in a more efficient or specific way than pure replacement.
+Chunks could even be moved, should the programmer add the ability to move the Earth--a programming habit whose prevalence has no clear evidence otherwise.
+This enforces expandability, which has more general positive (Cite etc) benefits.
+The \textbf{TimeMonitor}'s role is to ensure the map chunk generation can take place without a significant drop in frame rate.
+However, in order for the map to download at all, it must provide a minimum time space in case the machine is too slow to achieve it.
+A more advanced system may communicate with a hypothetical \textbf{FPSNormaliser} class, which records past processing times and reduces the framerate to a fixed, achievable value smoothest possible.
 
@@
 
-\begin{figure}
+\begin{figure}[H]
 	\includegraphics[width=0.7\linewidth]{Basic_Map_Streamer}

### 2018-03-23 draft.tex
Updated draft

@@
 This study aims to discover the extent of this problem, followed by proposals of architectural methods to implement systems that could mitigate the negative effects of an unreliable networking environment on the player experience.
-These can be described as QoE (Quality-of-Experience) improvements where the bandwidth is not improved, but instead coordinated where the player needs it most.
+These can be described as QoE (Quality-of-Experience) improvements where the bandwidth is not improved, but instead coordinated toward where the player needs it most.
 The programming language in question is C++, while the hypothetical game assumed a 3D environment wherein the player can manipulate the terrain and build meshes.
@@
 \begin{itemize}
-\item Exploration into the games market and its broadband speeds
-\item A brief introduction to player experience as a separate concept to content quality
-\item Existing strategies used by games for map streaming
+\item Exploration into the games market and its broadband speeds,
+\item A brief introduction to player experience as a separate concept to content quality, to gain an understanding of what should be prioritised,
+\item A priority list of assets and features of immediate interest to the player, and
 \item A proposed technical outline of a system that could facilitate level streaming with priority on player experience
@@
 There is a lack of definitive research on the impact of low broadband speed in map streaming.
-Instead, many studies such as [CITE] are focused on reducing and optimising traffic.
-This is helpful to developers, but lacks a point of reference or scale in terms of user experience, making the actual negative impacts of low bandwidth relatively uncertain.
-Until this is further understood, it could be beneficial to design a system around scalable broadband speed, such that it can guarantee an acceptably 'fun' quality even in extremely slow network conditions.
+A point of reference or scale in terms of user experience is therefore lacking, making the actual negative impacts of low bandwidth relatively uncertain.
+Until this is further understood, this essay will create the system around scalable broadband speed, aiming to guarantee an acceptably 'fun' quality even in extremely slow network conditions.
 However, the definition of 'fun' may be at odds with the industry's tendency to create games that are bigger, more detailed, and overall more data-heavy as they evolve \cite{graphicsvsexperience}.
@@
 Ryan and E.
-Deci's reputable research on human motivation \cite{motivation} (especially reputable in that they have disproportionately few people to cite, other than themselves, throughout their entire paper), which notes a human desire to explore, apply their skills, and learn from the feedback.
+Deci's reputable research on human motivation \cite{motivation} (noteworthy in that they have disproportionately few people to cite, other than themselves, throughout their entire paper), which notes a human desire to explore, apply their skills, and learn from the feedback.
 
@@
 	\item Player avatar
-	\item Textureless nearby level geometry
+	\item Untextured level geometry (nearby)
 	\item Interactive objects
@@
 	\item Object animations
+	\item Untextured level geometry (distant, low-detail)
 	\item Textures
-	\item Miscellaneous
+	\item Additional resources
 \end{itemize}
@@
 
-While they are described above, the specifics of animation, player model, object and texture streaming are beyond the scope of this essay.
-They could, however, very likely be adapted in a similar fashion.
-
-\section{Existing strategies}
-Locality-based streaming prioritisation already exists, and is used in open-world games to save memory and loading times.
-Engines such as the Unreal Engine provide World Composition, allowing the world to be partly loaded in small sections as the player approaches them \cite{unrealcomposition}.
-This is characterised by the loading of low-LOD (Level-of-Detail) versions of level components distant to the camera, and the high-LOD versions of close ones, all in a background process.
-This hides level details in the distance at the benefit of reducing the memory footprint, hiding loading times, and increasing the frame rate.
-The latter two in particular improves the player's perception of the game motion, and provides data to the player in a spatially-aware manner such that the player has immediate access to the areas closest to them.
-
-However, systems such as the above assume a higher data rate, and is typically sourced from a disk or hard disk.
+As an additional note, the specifics of animation, player model, object and texture streaming are beyond the scope of this essay, but are noted above for the reader's interest.
 
 \section{Proposal}
+The proposed system base, designed to prioritise the player experience, is illustrated below as two components, the client component and server component.
+These represent the game and the persistent online server, respectively.
+
+\subsection{Server system}
 \begin{figure}[H]
@@
 
-\textbf{Map}: Base map class.
+\begin{itemize}
+\item \textbf{Map}: The base map class.
 This is shared between the client and server.
-By partitioning the map into several 'chunks', separate pieces of it can be downloaded at a time, allowing the player to roam incomplete areas in short notice.
-\textbf{ChunkDestination}: A collection of \textbf{ChunkPacks} that may be implemented in multiple ways.
+By partitioning the map into several \textbf{MapChunks}, separate pieces of it can be downloaded at a time, allowing the player to roam incomplete areas in short notice.
+\item \textbf{MapChunk}: A portion of the map, with an index identifying the specific area of the level, and mesh information.
+The shape of a level chunk may vary across games, though in a building game, a uniform shape could be chosen such that there is consistency in the efficiency of building in one place over another.
+However, different games may use different strategies.
+For example, Sunset Overdrive (Insomniac Games, 2014) used hexagonal chunks by design and radial chunks in practice.
+This was optimised for the player's freedom of movement.
+However, an earlier game by the same company, Fuse (Insomniac Games, 2013), used rectangular chunks.
+This was optimised for linear hall-based gameplay \cite{overdrivegdc}.
+\item \textbf{ChunkDestination}: A collection of \textbf{ChunkPacks} that may be implemented in multiple ways.
 In this case, it is connected to \textbf{MessageSender} for immediate transmission.
 In more advanced cases, it could also, at a later stage, be implemented on the client machine as a map storage manager, which could cache map data on a local machine for later use.
-\textbf{MapUpstreamer}: The MapUpstreamer is effectively a \textit{Mediator} (described as a class that manages the collaboration between objects \cite{designpatterns}) technically owned by neither Map nor ChunkDestination.
+\item \textbf{MapUpstreamer}: The MapUpstreamer is effectively a \textit{Mediator} (described as a class that manages the collaboration between objects \cite{designpatterns}) technically owned by neither Map nor ChunkDestination.
 This method was chosen so that MapUpstreamer could, if necessary, operate independently in parallel to the other tasks.
 An observer pattern could reasonably be used, but this typically means synchronising the data on an immediate basis \cite{designpatterns}, something which is not necessary for the time-sensitive task of prioritising the data flow.
-\textbf{BandwidthMonitor}: To achieve data rate consistency as likely desired by players (see Section \ref{player}), this class monitors bandwidth over time and adjusts the data rate to the most consistent possible.
+\item \textbf{BandwidthMonitor}: To achieve data rate consistency as likely desired by players (see Section \ref{player}), this class monitors bandwidth over time and adjusts the data rate to the most consistent possible.
 The specifics of this system, better described as bandwidth-smoothing, are beyond the scope of this essay.
@@
 Rexford is recommended for further reading \cite{bandwidthsmoothing}.
-\textbf{ChunkPack}: This interface stores components of map data in various possible class-specific ways, and provides the function \textit{generateChunk} such that it can inject itself into the Map.
-An example of a subclass--\textbf{ChunkPack_Delta}--is demonstrated in Fig.
+\item \textbf{ChunkPack}: This interface stores components of map data in various possible class-specific ways, and provides the function \textit{generateChunk} such that it can inject itself into the Map.
+An example of a child class--\textbf{ChunkPack\_Delta}--is demonstrated in Fig.
 \ref{fig:clientsystem}.
+\end{itemize}
 
-The client side of the system we propose adopts the following hierarchy.
-
-\begin{figure}
+\subsection{Client system}
+\begin{figure}[H]
 	\centering
@@
 
-- The Map Streamer is a Listener class.
-This is to enable it to receive data from any input.
-A particular benefit of this approach is the ability to add different sources of data; for example, a local cache so that a player can return to a region without redownloading it.
-- The \textbf{ChunkPack} is an interface that can manipulate map chunks by inserting or removing data of its choice.
-By leaving this responsibility to the ChunkPack class, subclasses such as \textbf{ChunkPackDelta} can be created to manipulate the chunk data in a more efficient or specific way than pure replacement.
-Chunks could even be moved, should the programmer add the ability to move the Earth--a programming habit whose prevalence has no clear evidence otherwise.
-This enforces expandability, which has more general positive (Cite etc) benefits.
-The \textbf{TimeMonitor}'s role is to ensure the map chunk generation can take place without a significant drop in frame rate.
-However, in order for the map to download at all, it must provide a minimum time space in case the machine is too slow to achieve it.
-A more advanced system may communicate with a hypothetical \textbf{FPSNormaliser} class, which records past processing times and reduces the framerate to a fixed, achievable value smoothest possible.
+To support the independence of the background download process, which focuses on delivering data in a time-sensitive manner, this process is divided into two concurrent threads.
+This system comprises the following key classes, some of which are shared with the Server system:
 
-\section{Proposal}
-Due to an extreme lack of time and an extremely tired caffeine-free (and vitamin-deficient) author, a tacked-together solution is hastily proposed by this essay draft.
+\begin{itemize}
+\item \textbf{MapStreamer}: Responsible for collecting \textbf{ChunkPack}s and transferring them to the physical \textbf{Map} when it deems appropriate.
+This is another example of a Mediator.
+It relies on \textbf{TimeMonitor} to deliver the chunks smoothly.
+\item \textbf{TimeMonitor}: To ensure a smooth experience, this determines the amount of processing time to dedicate to chunk generation.
+This will be dependent on the game's other systems, and there are many factors to consider--such as used CPU time, network speed, the time it takes to generate an average chunk--to optimise this delivery time.
+\item \textbf{PlayerPerspective}: Responsible for informing the server where the player is in the map, so that the server can prioritise chunk delivery.
+\item \textbf{ChunkData\_Absolute}: Contains a complete map chunk at a certain level of detail.
+\item \textbf{ChunkData\_Delta}: Contains a map chunk in terms of changes that have occurred since an earlier version.
+This will be useful for player building events; for example, when a player constructs a wall.
+Contrast with rebuilding a mesh from scratch, this may return meshes that have been added or removed.
+\end{itemize}
 
-\begin{figure}[H]
-	\includegraphics[width=0.7\linewidth]{Basic_Map_Streamer}
-	\caption{A class diagram demonstrating a map streaming component connected to a basic Map class.}
-	\label{fig:simplesystem}
-\end{figure}
-
-The class diagram above depicts the client end of the map streaming game component.
-As some partly-downloaded data is not needed by the game component until it's complete, the process is divided into two concurrent, non-conflicting threads.
-Flow takes place in the following fashion:
+The class diagram above depicts the client end of the map streaming game component. Flow takes place in the following fashion:
 
@@
 
-A \textbf{level chunk} is defined as a region of space where static level geometry resides.
-The shape of a level chunk varies across games, though in a building game, a uniform shape could be chosen such that there are no notable inefficiencies to building in one place over another.
-Existing games use various strategies.
-For example, Sunset Overdrive (Insomniac Games, 2014) used hexagonal chunks by design and radial chunks in practice.
-This was optimised for the player's freedom of movement.
-However, an earlier game by the same company, Fuse (Insomniac Games, 2013), used rectangular chunks.
-This was optimised for linear hall-based gameplay.
-(CITE ALL THIS).
+\section{Conclusion}
+A system architecture focusing on delivering data to the player in a consistent \cite{qoelargestudy} \cite{lagragequits}, immediately interactive \cite{motivation}, and smooth \cite{graphicsvsexperience} \cite{lagragequits} way is proposed to optimise the player experience.
+While the environment was states to be C++, it is likely that this architecture could be applied to most modern object-oriented languages.
 
-The \textbf{level streamer} is responsible for collecting \textbf{Chunk Packs} (partial, LOD-specific level chunk data downloaded from the server) and transferring them to the physical \textbf{Map} when it deems appropriate.
-In the above example, it does so as soon as it is immediately downloaded.
-However, it may benefit the player experience to record the amount of processing time spent during recent frames, and responding by slowing the conversion process or postponing it.
-This could help preserve a high frame rate and provide a smoother experience.
-
-\section{Total Conclusion}
-There are further systems to be explored, citations to be added and slightly passive-aggressive expressions of frustration to be passive-aggressively expressed at the recent surge in workload that are beyond the scope of this essay draft, and the author's physical capabilities at 3am, let alone intellectual ones.
-The author as such has little idea what they are doing and submits this draft to the peer review with great hesitation and a tear in his eye.
+There is much room for further research and expansion, including but not limited to compression, variable chunk scaling, texture streaming, gameplay element streaming and sound streaming techniques.
+Furthermore, lack of research on the impact of bandwidth on building games leaves the true fidelity and efficiency requirements of this system to be quite undefined.
+This may be prompted by the development of more games of this type, which could facilitate further research into the connectivity requirements globally.
+Noting the popularity of Fortnite today, which now matches Minecraft \cite{minecraftnite}, this could be expected to happen in the near future.
 

### 2018-03-23 draft.tex
Probably the final revision

@@
 When downloading the game environment, low bandwidth increases download speeds and temporarily strips players of their autonomy, causing frustration.
-To aid this, Quality-of-Experience (QoE) techniques are proposed to mitigate the negative effects of limited bandwidth on the player experience, in a way that could be adapted to existing games with minimal difficulty.
+To aid this, Quality-of-Experience (QoE) techniques are proposed to mitigate the negative effects of limited bandwidth on the player experience, in a way that could be adapted to both new and existing games.
 \end{abstract}
@@
 
+This presents profitable opportunities for game developers to expand into new territories.
+However, a lag in technological advancement, particularly broadband, raises reasonable concerns of whether they can fully exploit these opportunities.
+The potential of games with building elements, such as the extremely popular (as suggested by Google Trends \cite{minecraftnite}) Minecraft and Fortnite, may be hindered by a lack of bandwidth necessary to download map resources at an acceptable speed.
+The slow download may frustrate players by rendering them unable to access areas of the game in a timely manner.
+
 \begin{figure}[H]
@@
 \end{figure}
-
-This presents profitable opportunities for game developers to expand into new territories.
-However, a lag in technological advancement, particularly broadband, raises reasonable concerns of whether they can fully exploit these opportunities.
-The potential of games where players can build the map, such as the extremely popular (as suggested by Google Trends \cite{minecraftnite}) Minecraft and Fortnite, may be hindered by a lack of bandwidth necessary to download player-built resources at an acceptable speed.
-A slow download of map resources may frustrate players by rendering them unable to access areas of the game in a timely manner.
 
@@
 \begin{itemize}
-\item Exploration into the games market and its broadband speeds,
-\item A brief introduction to player experience as a separate concept to content quality, to gain an understanding of what should be prioritised,
-\item A priority list of assets and features of immediate interest to the player, and
-\item A proposed technical outline of a system that could facilitate level streaming with priority on player experience
+\item Exploration into the games market and its broadband speeds
+\item A brief introduction to player experience as a separate concept to content quality, to gain an understanding of what should be prioritised
+\item A priority list of assets and features of immediate interest to the player
+\item A proposed technical outline of a system that could facilitate level streaming with priority on player experience.
 \end{itemize}
@@
 However, the definition of 'fun' may be at odds with the industry's tendency to create games that are bigger, more detailed, and overall more data-heavy as they evolve \cite{graphicsvsexperience}.
-This prompts an exploration into what is 'fun', in a term more aptly described as quality of experience.
+This prompts an exploration into what is 'fun', in a term more broadly described as quality-of-experience.
 
@@
 Studies have noted latency \cite{qossensitivity} \cite{lagragequits}, jitter and packet loss \cite{lagragequits} to all have negative impacts, sometimes leading to quitting altogether.
-While this is not widely studied in building games--but FPSes and MMOs--games such as the recent and highly popular \cite{topgames} Fortnite, which features both building elements and shooting gameplay, demonstrates that the two are not necessarily mutually exclusive.
+While this is not widely studied in building games (instead FPSes and MMOs), games such as \cite{topgames} Fortnite, which features both building and shooting gameplay, demonstrates that the two are not mutually exclusive.
 
@@
 Ryan and E.
-Deci's reputable research on human motivation \cite{motivation} (noteworthy in that they have disproportionately few people to cite, other than themselves, throughout their entire paper), which notes a human desire to explore, apply their skills, and learn from the feedback.
+Deci's research on human motivation \cite{motivation}.
+Whilst heartily citing themselves throughout their highly reputable paper, they observe a human desire to explore, apply their skills, and learn from the feedback.
 
 These findings place high importance on the motion, control, and abilities of the player rather than visual fidelity of the background.
-In map streaming, this translates to the importance of giving the player base landscapesto move within and toward.
+In map streaming, this translates to the importance of giving the player base landscapes to move within and toward.
 This places the level mesh, and perhaps interactive objects (noting the need for motion), above textures and sounds in the priority list.
@@
 \section{Quality of experience -- Priorities}
-To maximise player experience, a preliminary priority list is thus proposed:
+Based on the above findings, a preliminary priority list of asset downloads is proposed:
 
 \begin{itemize}
-	\item Player movement
-	\item Player avatar
-	\item Untextured level geometry (nearby)
-	\item Interactive objects
-	\item Player animations
-	\item Object animations
-	\item Untextured level geometry (distant, low-detail)
-	\item Textures
-	\item Additional resources
+	\item Player movement, to enable initial player autonomy and feedback \cite{motivation}
+	\item Player avatar, so they can see where they are, enhancing the above
+	\item Untextured level geometry (nearby), so the player has locations to visit
+	\item Interactive objects, for further feedback
+	\item Player animations and object animations to deliver the aesthetic \cite{graphicsvsexperience}
+	\item Untextured level geometry (distant, low-detail) for greater immersion
+	\item Textures and sounds to further enhance and complete the aesthetic
+	\item Additional resources.
 \end{itemize}
@@
 	\centering
-	\includegraphics[width=0.7\linewidth]{Server_Side_Streamer.png}
+	\includegraphics[width=1.0\linewidth]{Server_Side_Streamer.png}
 	\caption{Partial class diagram illustrating the potential map streaming server},
@@
 The shape of a level chunk may vary across games, though in a building game, a uniform shape could be chosen such that there is consistency in the efficiency of building in one place over another.
-However, different games may use different strategies.
+However, different games may use different strategies to best match their usage scenario.
 For example, Sunset Overdrive (Insomniac Games, 2014) used hexagonal chunks by design and radial chunks in practice.
-This was optimised for the player's freedom of movement.
+This was designed to facilitate the player's freedom of movement in every direction.
 However, an earlier game by the same company, Fuse (Insomniac Games, 2013), used rectangular chunks.
 This was optimised for linear hall-based gameplay \cite{overdrivegdc}.
+In a building game, the former would most likely be appropriate, as most games would likely take place in a wide virtual world so that the player can build large structures.
 \item \textbf{ChunkDestination}: A collection of \textbf{ChunkPacks} that may be implemented in multiple ways.
@@
 \item \textbf{ChunkPack}: This interface stores components of map data in various possible class-specific ways, and provides the function \textit{generateChunk} such that it can inject itself into the Map.
-An example of a child class--\textbf{ChunkPack\_Delta}--is demonstrated in Fig.
-\ref{fig:clientsystem}.
+See examples of subclasses in the following section.
 \end{itemize}
 
+\pagebreak
 \subsection{Client system}
@@
 	\centering
-	\includegraphics[width=0.7\linewidth]{Client_Side_Streamer.png}
+	\includegraphics[width=1.0\linewidth]{Client_Side_Streamer.png}
 	\caption{Partial class diagram illustrating the potential map streaming client},
@@
 \item \textbf{PlayerPerspective}: Responsible for informing the server where the player is in the map, so that the server can prioritise chunk delivery.
-\item \textbf{ChunkData\_Absolute}: Contains a complete map chunk at a certain level of detail.
+\item \textbf{ChunkData\_Absolute}: Contains a complete map chunk at a particular level of detail.
 \item \textbf{ChunkData\_Delta}: Contains a map chunk in terms of changes that have occurred since an earlier version.
 This will be useful for player building events; for example, when a player constructs a wall.
-Contrast with rebuilding a mesh from scratch, this may return meshes that have been added or removed.
+Contrast with rebuilding a mesh from scratch, this may return only meshes that have been added or removed, saving time and memory.
 \end{itemize}
-
-The class diagram above depicts the client end of the map streaming game component. Flow takes place in the following fashion:
-
-\begin{figure}[H]
-	\includegraphics[width=0.7\linewidth]{Basic_Flowchart.png}
-	\caption{Process flow of the two threads in parallel}
-	\label{fig:simpleflow}
-\end{figure}
-
-In this example, the client begins by sending the player and camera location as soon as possible.
-Sometimes these two variables may be different: it is important that both areas are prepared for the player's viewing.
-The server is assumed to prioritise the level chunk geometry from closest to furthest.
 
@@
 A system architecture focusing on delivering data to the player in a consistent \cite{qoelargestudy} \cite{lagragequits}, immediately interactive \cite{motivation}, and smooth \cite{graphicsvsexperience} \cite{lagragequits} way is proposed to optimise the player experience.
-While the environment was states to be C++, it is likely that this architecture could be applied to most modern object-oriented languages.
+While the development environment is implied to be C++, it is likely that this architecture could be applied to most modern object-oriented languages.
 
 There is much room for further research and expansion, including but not limited to compression, variable chunk scaling, texture streaming, gameplay element streaming and sound streaming techniques.
+A more detailed study could describe the particular methodology in the forms of flowcharts and psuedocode to help formulate the model.
 Furthermore, lack of research on the impact of bandwidth on building games leaves the true fidelity and efficiency requirements of this system to be quite undefined.
-This may be prompted by the development of more games of this type, which could facilitate further research into the connectivity requirements globally.
-Noting the popularity of Fortnite today, which now matches Minecraft \cite{minecraftnite}, this could be expected to happen in the near future.
+Observing the popularity of Fortnite today, which now matches Minecraft as of March 2018 \cite{minecraftnite}, the game industry is likely preparing to meet a new demand for games with buildling elements.
+Should this happen, their development should hope to drive this area of research forward.
 
